\sectionheader{Background and Related Work}
\label{sec:background}

\subsection{Petri Net Fundamentals}

Formally, a Petri Net is a tuple $(P, T, F, M_0)$ where:
\begin{itemize}
    \item $P$ is a finite set of places
    \item $T$ is a finite set of transitions
    \item $F \subseteq (P \times T) \cup (T \times P)$ is a set of arcs
    \item $M_0: P \rightarrow \mathbb{N}$ is the initial marking
\end{itemize}

The dynamics of Petri Nets are governed by the firing of transitions. A transition $t$ is enabled if each input place $p$ has at least as many tokens as the weight of the arc from $p$ to $t$. When a transition fires, it consumes tokens from its input places and produces tokens in its output places according to the weights of the corresponding arcs.

\subsection{Types of Petri Nets}

Several extensions to the basic Petri Net model have been proposed to enhance its modeling power:

\begin{itemize}
    \item \textbf{Colored Petri Nets} \cite{jensen1987coloured}: Extend the basic model by allowing tokens to have data values (colors).
    \item \textbf{Timed Petri Nets}: Incorporate time into the model, allowing for the analysis of temporal properties.
    \item \textbf{Stochastic Petri Nets}: Introduce probabilistic elements to model random behavior.
    \item \textbf{Hierarchical Petri Nets}: Allow for the decomposition of complex models into simpler submodels.
\end{itemize}

\subsection{Applications of Petri Nets}

Petri Nets have been applied to various domains, including:

\begin{itemize}
    \item \textbf{Workflow Management} \cite{van2000workflow}: Modeling and analyzing business processes.
    \item \textbf{Manufacturing Systems}: Modeling production lines and resource allocation.
    \item \textbf{Communication Protocols}: Analyzing the behavior of network protocols.
    \item \textbf{Software Design}: Modeling concurrent and distributed software systems.
\end{itemize}

\subsection{Related Work}

Petri Nets have been extended in various ways to model complex systems, but their integration with tensor algebra is relatively unexplored. Tensors provide a natural language for encoding multidimensional relationships and parallel computations, making them ideal for representing the state and evolution of Petri Nets with topological structure.

David Spivak’s work on polynomial functors provides a categorical framework for modeling data types and processes as compositional, tree-like structures. This perspective aligns naturally with Petri Nets, where transitions can be seen as functorial operations on collections of tokens, and places as objects in a category.

Digit-wise arithmetic is a canonical example of a computation that is both highly structured and parallelizable. Neural-symbolic systems have struggled to capture such computations in a transparent and generalizable way. By modeling digit-wise arithmetic as a tensorized Petri Net, we can make explicit the flow of information and the compositional structure of the computation.