\section{Future Directions and Discussion}

As AI advances, verifying and interpreting complex reasoning grows critical. The Absolute Zero Reasoner (AZR) paradigm shows that state-of-the-art reasoning can be achieved without human-curated data, relying solely on verifiable arithmetic operations as the foundation for truth. This aligns with the GASing philosophy: arithmetic, especially addition, provides a self-sufficient, axiomatically grounded framework for verifying correctness—independent of external supervision.

AZR's framework for abduction, deduction, and induction is ultimately reducible to arithmetic operations such as matrix dot products. By grounding all reasoning in arithmetic, AZR and GASing offer:
\begin{itemize}
    \item \textbf{Self-Sufficient Verification of Truth}: Arithmetic rules enable models to autonomously verify outputs, eliminating the need for human-provided labels.
    \item \textbf{Unified Framework for Reasoning}: Abduction, deduction, and induction can all be formalized as arithmetic operations, simplifying reasoning architectures.
    \item \textbf{Enhanced AI Interpretability and Trust}: Reducing complex reasoning to arithmetic steps makes the process auditable and explainable.
    \item \textbf{Bridging Symbolic and Sub-symbolic AI}: The arithmetic foundation enables integration between symbolic logic and neural computation.
    \item \textbf{Preparation for Autonomous AI}: Models can propose, solve, and verify tasks through self-play, guided by arithmetic truth, supporting scalable, autonomous learning.
\end{itemize}

The future of GASing and arithmetic-based reasoning lies in their ability to provide a universal, self-verifying substrate for all forms of inference, supporting scalable, interpretable, and trustworthy AI systems.
